---
title: "HW4"
author: "Xiayuanshan Gao, Jiahang Li, Yuhao Chen"
date: "2024-04-17"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: yes
    theme: cosmo 
    highlight: tango    
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE,cache = TRUE)
options(scipen=10000000)
library(RSocrata)
library(tidyverse)
library(kableExtra)
library(caret)
library(knitr) 
library(pscl)
library(plotROC)
library(pROC)
library(lubridate)
library(ggplot2)
library(gridExtra)

palette5 <- c("#eff3ff","#bdd7e7","#6baed6","#3182bd","#08519c")
palette4 <- c("#D2FBD4","#92BCAB","#527D82","#123F5A")
palette2 <- c("#6baed6","#08519c")
```

# Intro



```{r cars}
recidivism <- 
  read.socrata("https://data.ojp.usdoj.gov/Courts/NIJ-s-Recidivism-Challenge-Full-Dataset/ynf5-u8nk")
recidivism <- recidivism %>%
   mutate(recidivism_within_3years_numreic = case_when(
    recidivism_within_3years == "true" ~ 1,
    recidivism_within_3years == "false" ~ 0,
    TRUE ~ NA_integer_  # Handles any unexpected values
  ))


recidivism <- recidivism %>%
  drop_na(gender, race, age_at_release, supervision_level_first, education_level,
                   prison_offense, prison_years, violations, residence_puma,
                   supervision_risk_score_first, percent_days_employed, jobs_per_year,
                   avg_days_per_drugtest, drugtests_thc_positive, drugtests_meth_positive,
                   drugtests_cocaine_positive)

```


## Exploratory Analysis - Recidivism


```{r cache = TRUE, warning = FALSE, message = FALSE}
ggplot(data = recidivism) +
  geom_bar(aes(recidivism_within_3years,fill = recidivism_within_3years),width = 0.4) +
          scale_fill_manual(values = palette2,
                          name = 'Recidivism') +
  labs(x="Recidivism", y="Count", 
           title = "Non-recidivism vs.Recidivism\n",
       caption = 'Figure 2.1') +
      theme(legend.position = "none") 
```

## Categorical Variables

```{r cache = TRUE, warning = FALSE, message = FALSE,fig.width=10, fig.height=10}
cat_var <- recidivism %>%
  select(recidivism_within_3years, gender, race, age_at_release, supervision_level_first, education_level, prison_offense, prison_years, violations,prior_arrest_episodes_felony,prior_arrest_episodes_misd,delinquency_reports,residence_changes,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_conviction_episodes_7) %>%
  pivot_longer(cols = -recidivism_within_3years, names_to = "Variable", values_to = "value") %>%
  count(Variable, value, recidivism_within_3years) %>%
  ggplot(aes(x = value, y = n, fill = recidivism_within_3years)) + 
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~Variable, scales = "free", ncol = 4,
               labeller = labeller(Variable = c(
                 gender = "Gender",
                 race = "Race",
                 age_at_release = "Age at Release",
                 supervision_level_first = "Supervision Level",
                 education_level = "Education Level",
                 prison_offense = "Prison Offense",
                 prison_years = "Years in Prison",
                 violations = "Violations",
                 prior_arrest_episodes_felony = "Prior Arrest Episodes Felony",
                 prior_arrest_episodes_misd = "Prior Arrest Episodes Misd",
                 delinquency_reports="Delinquency Reports",
                 residence_changes = "Residence Changes",
                 prior_conviction_episodes_7 = "Prior Conviction Episodes:Gun Charges",
                 prior_conviction_episodes_6 = "Prior Conviction Episodes:Domestic Violence Charges",
                 prior_conviction_episodes_5 = "Prior Conviction Episodes:Probation Violation Charges",
                 prior_conviction_episodes_4 = "Prior Conviction Episodes:Drug",
                 prior_conviction_episodes_3 = "Prior Conviction Episodes:Property",
                 prior_conviction_episodes_2 = "Prior Conviction Episodes:Violence",
                 prior_conviction_episodes_1 = "Prior Conviction Episodes:Misdemeanor",
                 employment_exempt = "Employment Exempt"))) +
    scale_fill_manual(values = palette2, name = "Recidivism_within_3years") +
    labs(x = "Category", y = "Count",
         title = "Feature Associations with the Likelihood of Recidivism within 3 Years",
         subtitle = "Categorical Outcomes") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8), # Specify once
          plot.subtitle = element_text(size = 12, face = "italic"),
          plot.title = element_text(size = 18, face = "bold"),
          axis.text.y = element_text(size = 8),
          axis.title = element_text(size = 9),
          panel.background = element_blank(),
          panel.border = element_rect(colour = "grey", fill = NA, linewidth = 0.8))

print(cat_var)
```







## Chi square
```{r cache = TRUE, warning = FALSE, message = FALSE,fig.width=10, fig.height=10}
cat_var <- recidivism %>%
  select(recidivism_within_3years, gender, race, age_at_release, supervision_level_first, education_level, prison_offense, prison_years, violations)

chi_var <- c("gender", "race", "age_at_release", "supervision_level_first", "education_level", "prison_offense", "prison_years", "violations")
label_map <- c(gender = "Gender", race = "Race", age_at_release = "Age at Release",
               supervision_level_first = "Supervision Level First", education_level = "Education Level",
               prison_offense = "Prison Offense", prison_years = "Years in Prison", violations = "Violations")

chi_square_results <- data.frame(
  Variable = character(),
  Df = integer(),
  X_Squared = numeric(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)


for (var in chi_var) {
  contingency_table <- table(cat_var[[var]], cat_var$recidivism_within_3years)
  chi_square_test <- tryCatch({
    chisq.test(contingency_table)
  }, warning = function(w) {
    message("Warning for ", var, ": ", w$message)
    return(NULL)  # Return NULL if there's a warning
  }, error = function(e) {
    message("Error for ", var, ": ", e$message)
    return(NULL)  # Return NULL if there's an error
  })


  if (!is.null(chi_square_test)) {
    chi_square_results <- rbind(chi_square_results, data.frame(
      Variable = label_map[var],
      Df = chi_square_test$parameter,
      X_Squared = chi_square_test$statistic,
      P_Value = chi_square_test$p.value
    ))
  }
}
chi_square_results <- chi_square_results[, -1]

# Use the label_map you created earlier to set the row names to the descriptive labels
rownames(chi_square_results) <- label_map[chi_var]

# Generate the table without the first column of code variable names
kable(chi_square_results, col.names = c("Degrees of Freedom", "Chi-Squared", "P-Value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  footnote(general_title = "\n", general = "Table 1: Chi-square Test Results for Recidivism Variables")
```

## Numeric Variables
```{r cache = TRUE, warning = FALSE, message = FALSE,fig.width=10, fig.height=10}
numeric_var <- recidivism %>% 
  select(recidivism_within_3years, supervision_risk_score_first, percent_days_employed, jobs_per_year, avg_days_per_drugtest, drugtests_thc_positive, drugtests_meth_positive,drugtests_cocaine_positive)


numeric_var %>%
  gather(Variable, value, -recidivism_within_3years) %>%
    ggplot(aes(recidivism_within_3years, value, fill=recidivism_within_3years)) + 
      geom_bar(position = "dodge", stat = "summary", fun = "mean") + 
      facet_wrap(~Variable, scales = "free", ncol = 4, labeller= labeller(Variable = c(
                     `supervision_risk_score_first` = "Supervision Risk Score",
                     `percent_days_employed` = "Percent Days Employed",
                     `jobs_per_year` = "Jobs Per Year",
                     `avg_days_per_drugtest` = "Avg Days Per Drugtest",
                     `drugtests_thc_positive` = "Drugtests_THC_Positive",
                     `drugtests_meth_positive` = "Drugtests_Meth_Positive",
                     `drugtests_cocaine_positive`= "Drugtests_Cocaine_Positive"))) +
      scale_fill_manual(values = palette2) +
      labs(x="Recidivism_within_3years", y="Value", 
           title = "Feature Associations with the Likelihood of Recidivism",
           subtitle = "Numeric features") +
      theme(legend.position = "none") +
    theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 18, face = "bold"), 
        axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))
```





```{r cache = TRUE, warning = FALSE, message = FALSE, fig.height=10}
numeric_var %>%
    gather(Variable, value, -recidivism_within_3years) %>%
    ggplot() + 
    geom_density(aes(value, color=recidivism_within_3years), fill = "transparent") + 
   facet_wrap(~Variable, scales = "free", ncol = 4, labeller= labeller(Variable = c(
                     `residence_puma` = "Residence PUMA",
                     `supervision_risk_score_first` = "Supervision Risk Score",
                     `percent_days_employed` = "Percent Days Employed",
                     `jobs_per_year` = "Jobs Per Year",
                     `avg_days_per_drugtest` = "Avg Days Per Drugtest",
                     `drugtests_thc_positive` = "Drugtests_THC_Positive",
                     `drugtests_meth_positive` = "Drugtests_Meth_Positive",
                     `drugtests_cocaine_positive`= "Drugtests_Cocaine_Positive"))) +
  scale_color_manual(values = palette2) +
    labs(title = "Feature Distributions Based on Recidivism",
         subtitle = "Continous features") +
  theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 18, face = "bold"), 
        axis.text.x=element_text(size=9),
        axis.text.y=element_text(size=9), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))
```



## ANOVA TEST

```{r}
numeric_var <- recidivism %>%
  select(recidivism_within_3years, supervision_risk_score_first, percent_days_employed,
         jobs_per_year, avg_days_per_drugtest, drugtests_thc_positive,
         drugtests_meth_positive, drugtests_cocaine_positive) %>%
  na.omit()

# Define ANOVA variables and corresponding new readable names
anova_var <- c("supervision_risk_score_first", "percent_days_employed", "jobs_per_year",
               "avg_days_per_drugtest", "drugtests_thc_positive", "drugtests_meth_positive",
               "drugtests_cocaine_positive")

new_names <- c("Supervision Risk Score", "Percent Days Employed", "Jobs Per Year",
               "Avg Days Per Drugtest", "THC Positive Drugtests", "Meth Positive Drugtests",
               "Cocaine Positive Drugtests")

# Initialize the results data frame
anova_results <- data.frame(
  Variable = new_names,  # use new_names for the variable labels
  Df = integer(length(anova_var)),
  Sum_Sq = numeric(length(anova_var)),
  Mean_Sq = numeric(length(anova_var)),
  F_value = numeric(length(anova_var)),
  P_Value = numeric(length(anova_var)),
  stringsAsFactors = FALSE
)

# Run ANOVA for each variable
for (i in seq_along(anova_var)) {
  var <- anova_var[i]
  
  # Perform ANOVA
  anova_result <- aov(reformulate('recidivism_within_3years', response = var), data = numeric_var)
  
  # Extract the summary
  summary_data <- summary(anova_result)[[1]]
  
  # Fill the results data frame
  anova_results[i, "Df"] <- summary_data["Df"][1]
  anova_results[i, "Sum_Sq"] <- summary_data["Sum Sq"][1]
  anova_results[i, "Mean_Sq"] <- summary_data["Mean Sq"][1]
  anova_results[i, "F_value"] <- summary_data["F value"][1]
  anova_results[i, "P_Value"] <- summary_data["Pr(>F)"][1]
}

# Output the results
anova_results %>%
  kable(col.names = c("Variable", "Degrees of Freedom", "Sum Sq", "Mean Sq", "F value", "P-Value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  footnote(general_title = "\n", general = "Table 2: ANOVA Results for Numeric Variables")
```








# Logistic Regression




We then fit a logistic regression containing all variables in our dataset. To do so, we split our data randomly into 65% of training dataset and 35% of testing dataset. Some of our variables, such as Lien and Education, have single occurrence and therefore needs to be put into the training set to avoid new levels appearing in the testing set. This model shows that some indicators are significantly correlated with credit acceptance with p < 0.05. This include age, some categories of job, ways of contact, month last contacted, previous campaign outcome, unemployment rate, consumer price and confidence indices. The majority of predictors are not significant.

## Feature Engineering
```{r cache = TRUE, warning = FALSE, message = FALSE,fig.width=10, fig.height=10}
# age
recidivism <-recidivism %>%
  mutate(age = case_when(
    age_at_release == "18-22"  ~ "Gen Z",
    age_at_release =="38-42" |age_at_release == "43-47" |age_at_release == "48 or older"  ~ "Gen X",
    age_at_release == "23-27" |age_at_release == "28-32" |age_at_release == "33-37" ~ "Millenials"))

recidivism <-recidivism %>%
  mutate(jobnumber = cut(jobs_per_year,
                        breaks = c(-Inf, 1, 5, Inf),  # Define intervals, including all numbers
                        labels = c("0-1", "2-5", "6-8"),  # Labels for each interval
                        include.lowest = TRUE) )

recidivism <-recidivism %>%
  mutate(employmentstates = cut(percent_days_employed,
                        breaks = c(-Inf, 0, 0.99, Inf),  # Define intervals, including all numbers
                        labels = c("Unemployed", "Parttime emplyed", "Employed"),  # Labels for each interval
                        include.lowest = TRUE) )

```



```{r}
set.seed(479)
trainIndex <- createDataPartition(y = paste(recidivism$race) , p = .65,
                                  list = FALSE,
                                  times = 1)
```


```{r}
recidivismTrain <- recidivism[ trainIndex,]
recidivismTest  <- recidivism[-trainIndex,]

recidivismModel <- glm(recidivism_within_3years_numreic ~ .,
                  data=recidivismTrain %>% 
                    dplyr::select(recidivism_within_3years_numreic,gender,race,age_at_release,dependents, supervision_level_first,education_level,prison_years,prison_offense,violations,supervision_risk_score_first, percent_days_employed,jobs_per_year,drugtests_thc_positive,drugtests_cocaine_positive, drugtests_meth_positive,gang_affiliated,prior_arrest_episodes_felony,prior_arrest_episodes_misd,prior_arrest_episodes_property,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_revocations_probation,prior_revocations_parole,delinquency_reports,residence_changes,violations_instruction), family="binomial" (link="logit"))

recidivism_sum <- summary(recidivismModel)
coefficients_table <- as.data.frame(recidivism_sum$coefficients)

coefficients_table$significance <- ifelse(coefficients_table$`Pr(>|z|)` < 0.001, '***',
                                         ifelse(coefficients_table$`Pr(>|z|)` < 0.01, '**',
                                                ifelse(coefficients_table$`Pr(>|z|)` < 0.05, '*',
                                                       ifelse(coefficients_table$`Pr(>|z|)` < 0.1, '.', ''))))

coefficients_table$p_value <- paste0(round(coefficients_table$`Pr(>|z|)`, digits = 3), coefficients_table$significance)

coefficients_table %>%
  select(-significance, -`Pr(>|z|)`) %>% 
  kable(align = "r") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  footnote(general_title = "\n", general = "Table 3")
```

```{r}
pR2(recidivismModel) 
```

```{r}
testProbs <- data.frame(Outcome = as.factor(recidivismTest$recidivism_within_3years_numreic), Probs = predict(recidivismModel, recidivismTest, type= "response"))
head(testProbs)
```

Visually we now can begin comparing the predictions of recidivism. A negative or 0 value means that there is no recidivism, while a positive or 1 value means that there is recidivism. Strong models will have a peak closer to 0 for the negatives (no recidivism), and a peak closer to 1 for the positives (recidivism). In the following graph we can see that our model is better at prediciting the positives rather than the negatives.




```{r}
ggplot(testProbs, aes(x = Probs, fill = as.factor(Outcome))) + 
  geom_density() +
  facet_grid(Outcome ~ .) +
  scale_fill_manual(values = palette2) + xlim(0, 1) +
  labs(x = "Recidivism", y = "Density of probabilities",
       title = "Distribution of predicted probabilities by observed outcome") + theme(strip.text.x = element_text(size = 18),
        legend.position = "none")
```


## ROC Curve

We computed the Area Under the Receiver Operating Characteristic Curve (AUC-ROC) metric to further evaluate the performance of our improved model. The AUC value ranges from 0 to 1, where 0.5 means a model with no discriminatory power (it performs as well as random chance), and 1 means a perfect model (it perfectly distinguishes between positive and negative cases).Generally, an AUC between 0.7 and 0.8 is considered acceptable and indicates a fair to good performance of the model. While there is room for improvement, a score of 0.74 suggests that the model is capturing important patterns in the data and making predictions better than random guessing.

```{r}
ggplot(testProbs, aes(d = as.numeric(testProbs$Outcome), m = Probs)) +
  geom_roc(n.cuts = 50, labels = FALSE, colour = "#FE9900") +
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = 'grey') +
  labs(title = "Figure 5: ROC Curve - Recidivism")
```

```{r}
pROC::auc(testProbs$Outcome, testProbs$Probs) # AUC: .73
```

## Confusion Matrix

The Confusion Matrix shows the number of observed instances of using the tax credit that are predicted as such. Each entry in the matrix provides a different comparison between observed and predicted, given the 30% threshold.

```{r}
testProbs <- 
  testProbs %>%
  mutate(predOutcome  = as.factor(ifelse(testProbs$Probs > 0.3 , 1, 0)))

caret::confusionMatrix(testProbs$predOutcome, testProbs$Outcome, 
                       positive = "1")
```




## Cross Validation 

Next, we performed cross validation test for our model. The trainControl parameter is set to run 100 k-folds and to output predicted probabilities, classProbs, for two classes, recidivism and non-recidivism. Additional parameters output AUC (the train function refers to this as ‘ROC’) and confusion metrics for each fold. The three metrics in the cvFit output are for mean AUC, Sensitivity, and Specificity across all 100 folds. Here, the ROC for our model is 0.775, the Sensitivity is 0.8, while the Specificity is 0.6.





```{r}
ctrl <- trainControl(method = "cv", number = 100, classProbs = TRUE, summaryFunction = twoClassSummary)

# Fit the model
cvFit <- train(recidivism_within_3years ~ .,
               data = recidivism %>% 
                 select(recidivism_within_3years, gender, race, age_at_release, dependents, supervision_level_first, education_level, prison_years, prison_offense, violations, supervision_risk_score_first, percent_days_employed, jobs_per_year, drugtests_thc_positive, drugtests_cocaine_positive, drugtests_meth_positive, gang_affiliated, prior_arrest_episodes_felony, prior_arrest_episodes_misd, prior_arrest_episodes_property, prior_conviction_episodes_6, prior_conviction_episodes_5, prior_conviction_episodes_4, prior_conviction_episodes_3, prior_conviction_episodes_2, prior_conviction_episodes_1, prior_revocations_probation, prior_revocations_parole, delinquency_reports, residence_changes, violations_instruction),
               method = "glm", family = "binomial",
               metric = "ROC", trControl = ctrl)

# Output the fit
print(cvFit)
```
The figure below plots the distribution of AUC, Sensitivity, and Specificity across the 100 folds. The tighter each distribution is to its mean, the more generalizable the model. Both models generalizes well to sensitivity – the rate it correctly predicts credit acceptance, instead of specificity. It seems our would-be decision-making tool is inconsistent in how it predicts the credit acceptance. This inconsistency will have a direct effect on the marketing and resource allocation process should this algorithm be put into production.

```{r}
dplyr::select(cvFit$resample, -Resample) %>%
  gather(metric, value) %>%
  left_join(gather(cvFit$results[2:4], metric, mean)) %>%
  ggplot(aes(value)) + 
    geom_histogram(bins=35, fill = "#6baed6") +
    facet_wrap(~metric) +
    geom_vline(aes(xintercept = mean), colour = "#08519c", linetype = 3, size = 1.5) +
    scale_x_continuous(limits = c(0, 1)) +
    labs(x="Goodness of Fit", y="Count", title="CV Goodness of Fit Metrics",
         subtitle = "Across-fold mean reprented as dotted lines")

```

# Generalizability

## Racial Fairness

```{r}
recidivism %>%
    group_by(recidivism_within_3years, race) %>%
    summarize(n = n()) %>%
    mutate(freq = n / sum(n)) %>% filter(recidivism_within_3years == "true") %>%
    ggplot(aes(reorder(race, -freq), freq)) +
    geom_bar(stat = "identity", position = "dodge", fill = palette2) +
    labs(title = "Recidivism rate by race",
         y = "Rate", x = "Race") +
    theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1))
```



```{r}
reg.noRace <- glm(recidivism_within_3years_numreic ~ ., data = 
                    recidivismTrain %>% dplyr::select(recidivism_within_3years_numreic,gender,age_at_release,dependents, supervision_level_first,education_level,prison_years,prison_offense,violations,supervision_risk_score_first, percent_days_employed,jobs_per_year,drugtests_thc_positive,drugtests_cocaine_positive, drugtests_meth_positive,gang_affiliated,prior_arrest_episodes_felony,prior_arrest_episodes_misd,prior_arrest_episodes_property,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_revocations_probation,prior_revocations_parole,delinquency_reports,residence_changes,violations_instruction),
                family = "binomial"(link = "logit"))

reg.withRace <- glm(recidivism_within_3years_numreic ~ ., data = 
                      recidivismTrain %>% dplyr::select(recidivism_within_3years_numreic,gender,race,age_at_release,dependents, supervision_level_first,education_level,prison_years,prison_offense,violations,supervision_risk_score_first, percent_days_employed,jobs_per_year,drugtests_thc_positive,drugtests_cocaine_positive, drugtests_meth_positive,gang_affiliated,prior_arrest_episodes_felony,prior_arrest_episodes_misd,prior_arrest_episodes_property,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_revocations_probation,prior_revocations_parole,delinquency_reports,residence_changes,violations_instruction),
                family = "binomial"(link = "logit"))

```


```{r}
testProbs_race <- 
  data.frame(class = recidivismTest$recidivism_within_3years_numreic,
             probs = predict(reg.noRace, recidivismTest, type = "response"),
             Race = recidivismTest$race)
```

```{r}
mutate(testProbs_race, predClass = ifelse(probs >= .5, 1, 0)) %>%
  group_by(Race) %>%
  summarize(Observed.recidivism = sum(class) / n(),
            Predicted.recidivism = sum(predClass) / n()) %>%
  gather(Variable, Value, -Race) %>%
  ggplot(aes(Race, Value)) +
    geom_bar(aes(fill = Race), position="dodge", stat="identity") +
    scale_fill_manual(values = palette2) +
    facet_wrap(~Variable) +
    labs(title = "Observed and predicted recidivism", x = "Race", y = "Rate") +
    theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1))
```
```{r}
iterateThresholds <- function(data, observedClass, predictedProbs, group) {
  observedClass <- enquo(observedClass)
  predictedProbs <- enquo(predictedProbs)
  group <- enquo(group)
  x = .01
  all_prediction <- data.frame()
  
  if (missing(group)) {
  
    while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x,2))
    
    all_prediction <- rbind(all_prediction,this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
  else if (!missing(group)) { 
   while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      group_by(!!group) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x, 2))
    
    all_prediction <- rbind(all_prediction, this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
}
```

```{r}
testProbs_race.thresholds <- 
  iterateThresholds(data=testProbs_race, observedClass = class, 
                    predictedProbs = probs, group = Race)

filter(testProbs_race.thresholds, Threshold == .5)  %>%
  dplyr::select(Accuracy, Race, starts_with("Rate")) %>%
  gather(Variable, Value, -Race) %>%
    ggplot(aes(Variable, Value, fill = Race)) +
      geom_bar(aes(fill = Race), position = "dodge", stat = "identity") +
      scale_fill_manual(values = palette2) +
      labs(title="Confusion matrix rates by race",
           subtitle = "50% threshold", x = "Outcome",y = "Rate") +
      theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1)) 
```
```{r}
# Calculate the ROC curve and AUC for each race
roc_list <- lapply(split(testProbs_race, testProbs_race$Race), function(subdata) {
  roc(response = subdata$class, predictor = subdata$probs)
})

# Create an empty ggplot object with the correct aesthetics
ggroc_combined <- ggplot() + theme_minimal() + 
  labs(title = "ROC Curves by race", x = "False Positive Rate", y = "True Positive Rate") +
  geom_abline(slope=1, intercept=0, linetype="dashed", color = "grey")

# Calculate the position for AUC annotations
max_specificity <- max(sapply(roc_list, function(x) max(1 - x$specificities)))
min_sensitivity <- min(sapply(roc_list, function(x) min(x$sensitivities)))

# Adding ROC curves to the plot with annotations for AUC
for (i in seq_along(roc_list)) {
  roc_data <- roc_list[[i]]
  auc_value <- auc(roc_data)
  ggroc_combined <- ggroc_combined + 
    geom_line(data = data.frame(fp = 1 - roc_data$specificities, tp = roc_data$sensitivities),
              aes(x = fp, y = tp), color = palette2[i], size = 0.5) +
    geom_point(data = data.frame(fp = 1 - roc_data$specificities, tp = roc_data$sensitivities),
               aes(x = fp, y = tp), color = palette2[i], size = 0.3) +
    annotate("text", x = max_specificity - 0.1, y = min_sensitivity + 0.05 * (length(roc_list) - i + 1), 
             label = paste(names(roc_list)[i], "AUC =", round(auc_value, 3)), 
             hjust = 1, color = palette2[i], size = 3.5)
}

print(ggroc_combined)
```


## Racial-Gender Fairness 

```{r}

recidivism <- recidivism %>%
  mutate(gender = case_when(
    gender == "F" ~ "Female",
    gender == "M" ~ "Male",
    TRUE ~ as.character(gender)
  ))

recidivism <- recidivism %>%
  mutate(race = case_when(
    race == "BLACK" ~ "Black",
    race == "WHITE" ~ "White",
    TRUE ~ as.character(race)
  ))

recidivism$race_gender <- interaction(recidivism$race, recidivism$gender, sep = "_")

recidivism %>%
    group_by(recidivism_within_3years, race_gender) %>%
    summarize(n = n()) %>%
    mutate(freq = n / sum(n)) %>% filter(recidivism_within_3years == "true") %>%
    ggplot(aes(reorder(race_gender, -freq), freq)) +
    geom_bar(stat = "identity", position = "dodge", fill = palette4) +
    labs(title = "Recidivism rate by race and gender",
         y = "Rate", x = "Race_Gender") +
    theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1))
```

```{r}
recidivismTrain <- recidivism[ trainIndex,]
recidivismTest  <- recidivism[-trainIndex,]
```



```{r}
reg.norace_gender <- glm(recidivism_within_3years_numreic ~ ., data = 
                    recidivismTrain %>% dplyr::select(recidivism_within_3years_numreic,gender,race,age_at_release,dependents, supervision_level_first,education_level,prison_years,prison_offense,violations,supervision_risk_score_first, percent_days_employed,jobs_per_year,drugtests_thc_positive,drugtests_cocaine_positive, drugtests_meth_positive,gang_affiliated,prior_arrest_episodes_felony,prior_arrest_episodes_misd,prior_arrest_episodes_property,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_revocations_probation,prior_revocations_parole,delinquency_reports,residence_changes,violations_instruction),
                family = "binomial"(link = "logit"))

reg.withrace_gender <- glm(recidivism_within_3years_numreic ~ ., data = 
                      recidivismTrain %>% dplyr::select(recidivism_within_3years_numreic,gender,race,age_at_release,dependents, supervision_level_first,education_level,prison_years,prison_offense,violations,supervision_risk_score_first, percent_days_employed,jobs_per_year,drugtests_thc_positive,drugtests_cocaine_positive, drugtests_meth_positive,gang_affiliated,prior_arrest_episodes_felony,prior_arrest_episodes_misd,prior_arrest_episodes_property,prior_conviction_episodes_6,prior_conviction_episodes_5,prior_conviction_episodes_4,prior_conviction_episodes_3,prior_conviction_episodes_2,prior_conviction_episodes_1,prior_revocations_probation,prior_revocations_parole,delinquency_reports,residence_changes,violations_instruction),
                family = "binomial"(link = "logit"))

```

```{r}
testProbs_race_gender <- 
  data.frame(class = recidivismTest$recidivism_within_3years_numreic,
             probs = predict(reg.norace_gender, recidivismTest, type = "response"),
             Race_Gender = recidivismTest$race_gender)
```

```{r}
mutate(testProbs_race_gender, predClass = ifelse(probs >= .5, 1, 0)) %>%
  group_by(Race_Gender) %>%
  summarize(Observed.recidivism = sum(class) / n(),
            Predicted.recidivism = sum(predClass) / n()) %>%
  gather(Variable, Value, -Race_Gender) %>%
  ggplot(aes(Race_Gender, Value)) +
    geom_bar(aes(fill = Race_Gender), position="dodge", stat="identity") +
    scale_fill_manual(values = palette4) +
    facet_wrap(~Variable) +
    labs(title = "Observed and predicted recidivism", x = "Race_Gender", y = "Rate") +
    theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1))
```

```{r}
iterateThresholds <- function(data, observedClass, predictedProbs, group) {
  observedClass <- enquo(observedClass)
  predictedProbs <- enquo(predictedProbs)
  group <- enquo(group)
  x = .01
  all_prediction <- data.frame()
  
  if (missing(group)) {
  
    while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x,2))
    
    all_prediction <- rbind(all_prediction,this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
  else if (!missing(group)) { 
   while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      group_by(!!group) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x, 2))
    
    all_prediction <- rbind(all_prediction, this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
}
```

```{r}
testProbs_Race_Gender.thresholds <- 
  iterateThresholds(data=testProbs_race_gender, observedClass = class, 
                    predictedProbs = probs, group = Race_Gender)

filter(testProbs_Race_Gender.thresholds, Threshold == .5)  %>%
  dplyr::select(Accuracy, Race_Gender, starts_with("Rate")) %>%
  gather(Variable, Value, -Race_Gender) %>%
    ggplot(aes(Variable, Value, fill = Race_Gender)) +
      geom_bar(aes(fill = Race_Gender), position = "dodge", stat = "identity") +
      scale_fill_manual(values = palette4) +
      labs(title="Confusion matrix rates by race and gender",
           subtitle = "50% threshold", x = "Outcome",y = "Rate") +
      theme_minimal() + theme(axis.text.x = element_text(angle = 0, hjust = 1)) 
```
```{r}
# Calculate the ROC curve and AUC for each race
roc_list <- lapply(split(testProbs_race_gender, testProbs_race_gender$Race_Gender), function(subdata) {
  roc(response = subdata$class, predictor = subdata$probs)
})

# Create an empty ggplot object with the correct aesthetics
ggroc_combined <- ggplot() + theme_minimal() + 
  labs(title = "ROC Curves by race", x = "False Positive Rate", y = "True Positive Rate") +
  geom_abline(slope=1, intercept=0, linetype="dashed", color = "grey")

# Calculate the position for AUC annotations
max_specificity <- max(sapply(roc_list, function(x) max(1 - x$specificities)))
min_sensitivity <- min(sapply(roc_list, function(x) min(x$sensitivities)))

# Adding ROC curves to the plot with annotations for AUC
for (i in seq_along(roc_list)) {
  roc_data <- roc_list[[i]]
  auc_value <- auc(roc_data)
  ggroc_combined <- ggroc_combined + 
    geom_line(data = data.frame(fp = 1 - roc_data$specificities, tp = roc_data$sensitivities),
              aes(x = fp, y = tp), color = palette4[i], size = 0.5) +
    geom_point(data = data.frame(fp = 1 - roc_data$specificities, tp = roc_data$sensitivities),
               aes(x = fp, y = tp), color = palette4[i], size = 0.3) +
    annotate("text", x = max_specificity - 0.1, y = min_sensitivity + 0.05 * (length(roc_list) - i + 1), 
             label = paste(names(roc_list)[i], "AUC =", round(auc_value, 3)), 
             hjust = 1, color = palette4[i], size = 3.5)
}

print(ggroc_combined)
```


#  Cost-Benefit Analysis 

```{r}
cost_benefit_table <-
   testProbs %>%
      count(predOutcome, Outcome) %>%
      summarize(True_Negative = sum(n[predOutcome==0 & Outcome==0]),
                True_Positive = sum(n[predOutcome==1 & Outcome==1]),
                False_Negative = sum(n[predOutcome==0 & Outcome==1]),
                False_Positive = sum(n[predOutcome==1 & Outcome==0])) %>%
       gather(Variable, Count) %>%
       mutate(Revenue =
               ifelse(Variable == "True_Negative", Count * 0,
               ifelse(Variable == "True_Positive",((.35 - .1) * Count),
               ifelse(Variable == "False_Negative", (-0.35) * Count,
               ifelse(Variable == "False_Positive", (-0.1) * Count, 0))))) %>%
    bind_cols(data.frame(Description = c(
              "We correctly predicted no recidivism",
              "We correctly predicted a recidivism",
              "We predicted no recidivism but the crimer recidivate",
              "We predicted a recidivism but the crimer did not recidivate")))

kable(cost_benefit_table,
       caption = "Cost/Benefit Table") %>% kable_styling()
```

```{r}
iterateThresholds <- function(data) {
  x = .01
  all_prediction <- data.frame()
  while (x <= 1) {
  
  this_prediction <-
      testProbs %>%
      mutate(predOutcome = ifelse(Probs > x, 1, 0)) %>%
      count(predOutcome, Outcome) %>%
      summarize(True_Negative = sum(n[predOutcome==0 & Outcome==0]),
                True_Positive = sum(n[predOutcome==1 & Outcome==1]),
                False_Negative = sum(n[predOutcome==0 & Outcome==1]),
                False_Positive = sum(n[predOutcome==1 & Outcome==0])) %>%
     gather(Variable, Count) %>%
     mutate(Revenue =
               ifelse(Variable == "True_Negative", Count * 0,
               ifelse(Variable == "True_Positive",((.35 - .1) * Count),
               ifelse(Variable == "False_Negative", (-0.35) * Count,
               ifelse(Variable == "False_Positive", (-0.1) * Count, 0)))),
            Threshold = x)
  
  all_prediction <- rbind(all_prediction, this_prediction)
  x <- x + .01
  }
return(all_prediction)
}

```

```{r}
whichThreshold <- iterateThresholds(testProbs)

whichThreshold_revenue <- 
whichThreshold %>% 
    group_by(Threshold) %>% 
    summarize(Revenue = sum(Revenue))

whichThreshold %>%
  ggplot(.,aes(Threshold, Revenue, colour = Variable)) +
  geom_point() +
  scale_colour_manual(values = palette5[c(5, 1:3)]) +    
  labs(title = "Figure 9: Revenue by confusion matrix type and threshold",
       y = "Revenue") +
  theme_minimal() +
  guides(colour=guide_legend(title = "Confusion Matrix")) 
```


```{r}
whichThreshold_revenue <- 
  whichThreshold %>% 
    mutate(TookCredit = ifelse(Variable == "True_Positive", (Count * .25),
                         ifelse(Variable == "False_Negative", Count, 0))) %>%
  group_by(Threshold) %>% 
    summarize(Total_Revenue = sum(Revenue),
              Total_Count_Of_Credits = sum(TookCredit))

# Revenue 
grid.arrange(ncol = 1,
ggplot(whichThreshold_revenue)+ 
  geom_line(aes(x = Threshold, y = Total_Revenue))+
  geom_vline(xintercept =  pull(arrange(whichThreshold_revenue, -Total_Revenue)[1,1]))+
    labs(title = "Figure 10: Total Revenues By Threshold",
         subtitle = "Vertical Line Denotes Optimal Threshold"),
# Credits
ggplot(whichThreshold_revenue)+ 
  geom_line(aes(x = Threshold, y = Total_Count_Of_Credits))+
  geom_vline(xintercept =  pull(arrange(whichThreshold_revenue, -Total_Count_Of_Credits)[1,1]))+
    labs(title = "Figure 11: Total Count of Credits By Threshold",
         subtitle = "Vertical Line Denotes Optimal Threshold"))
```

#  Conclusion






